Bayesian Nonparametric System Reliability\\ using Sets of Priors

### 2016-01-14 ijar-npb-ms.tex
Section 5 language tweaks -- feel free to roll back any/all of them!

As was shown by \citet{2009:WalterAugustin}, %Walter \& Augustin (2009),
we can have both tractability and meaningful reaction to prior-data conflict
by using sets of priors $\MktZ$ produced by parameter sets $\PktZ = [\nktzl, \nktzu] \times [\yktzl, \yktzu]$. 
(A detailed discussion of different choices for $\PktZ$ is given in \citet[\S 3.1]{2013:diss-gw}.)
In our model, each prior parameter pair $(\nktz, \yktz) \in \PktZ$
corresponds to a Beta prior, thus $\MktZ$ is a set of Beta priors.
The set of posteriors $\MktN$ is obtained by updating each prior in $\MktZ$ according to Bayes' Rule.
This element-by-element updating may seem ad-hoc, but can be rigorously justified
as ensuring coherence \citep[\S 2.5]{1991:walley}, and was termed ``Generalized Bayes' Rule'' by \citet[\S 6.4]{1991:walley}.
Due to conjugacy, $\MktN$ is a set of Beta distributions with parameters $(\nktn, \yktn)$
obtained from updating $(\nktz, \yktz) \in \PktZ$ according to \eqref{eq:nyupdate},
and so can be described efficiently through the set of updated parameters
\begin{align}
\PktN &= \Big\{ (\nktn, \yktn) \mid (\nktz, \yktz) \in \PktZ = [\nktzl, \nktzu] \times [\yktzl, \yktzu] \Big\}\,.
\label{eq:paramsets}
\end{align}
Examples for parameter sets $\PktZ$ and $\PktN$ as in \eqref{eq:paramsets} are depicted in Figure~\ref{fig:paramsets}.
Such rectangular prior parameter sets $\PktZ$ have been shown
to balance desirable model properties and ease of elicitation quite well
(see \citet[pp.~123f]{2013:diss-gw} or \citet{Troffaes2013a-short}).
%
For each component type $k$ and time point $t$,
one has to choose only the four parameters $\nktzl, \nktzu, \yktzl, \yktzu$
(so $4 \times |{\cal T}|$ parameters are needed to define the set of prior distributions characterizing
prior knowledge on the survival function of components of type $k$);
notwithstanding this, the model has favourable inference properties.

@@

Central to these inference properties is that
while the prior parameter set $\PktZ$ is rectangular,
the posterior parameter set $\PktN$ is not rectangular anymore;
indeed, the shape of $\PktN$ depends on the presence of prior-data conflict,
which is naturally operationalized as $s^k_t/n_k \not\in [\yktzl, \yktzu]$:
prior-data conflict is present when, at time $t$, the observed fraction of functioning components
is outside its expected range.

In absence of prior-data conflict, 
$\PktN$ shrinks in the $\ykt$ dimension;
how much it shrinks depends on $\nktz \in [\nktzl, \nktzu]$,
leading to the so-called spotlight shape depicted in Figure~\ref{fig:paramsets} (left).
As $\yktn$ gives the posterior expectation for the functioning probability $p_t^k$,
shorter $\yktn$ intervals mean more precise knowledge about $p_t^k$.
Also, the variance interval for $p_t^k$ will shorten and shift towards zero,
as the Beta distributions in $\MktN$ will be more pointed
due to the increase of $\nktz$ to $\nktn$.

In case of prior-data conflict ($s^k_t/n_k \not\in [\yktzl, \yktzu]$),
$\PktN$ has instead the so-called banana shape,
arising from $\yktn$ intervals being shifted closer towards $s^k_t/n_k$
for lower $\nktn$ values than for higher $\nktn$ values, see Figure~\ref{fig:paramsets} (right).
Overall, this results in a wider $\yktn$ interval as compared to the no conflict case, 
reflecting the extra uncertainty due to prior-data conflict,
thus making more cautious probability statements about $p_t^k$.

Based on these shapes and \eqref{eq:nyupdate}, it is possible to deduce the following expressions
for the lower and upper bounds of $\yktn$:
\begin{align}
\begin{aligned}
\min_{\PktN} \yktn &=
 \begin{cases}
 %\dfrac{\nktzu \yktzl + s^k_t}{\nktzu + n_k} & \text{if } s^k_t/n_k \ge \yktzl \\
 \big(\nktzu \yktzl + s^k_t\big) / \big(\nktzu + n_k\big) & \text{if } s^k_t/n_k \ge \yktzl \\
 %\dfrac{\nktzl \yktzl + s^k_t}{\nktzl + n_k} & \text{if } s^k_t/n_k <   \yktzl
 \big(\nktzl \yktzl + s^k_t\big) / \big(\nktzl + n_k\big) & \text{if } s^k_t/n_k <   \yktzl
 \end{cases}\,,\\
\max_{\PktN} \yktn &=
 \begin{cases}
 %\dfrac{\nktzu \yktzu + s^k_t}{\nktzu + n_k} & \text{if } s^k_t/n_k \le \yktzu \\
 \big(\nktzu \yktzu + s^k_t\big) / \big(\nktzu + n_k\big) & \text{if } s^k_t/n_k \le \yktzu \\
 %\dfrac{\nktzl \yktzu + s^k_t}{\nktzl + n_k} & \text{if } s^k_t/n_k >   \yktzu
 \big(\nktzl \yktzu + s^k_t\big) / \big(\nktzl + n_k\big) & \text{if } s^k_t/n_k >   \yktzu
 \end{cases}\,.
\end{aligned}
\label{eq:ysetupdate}
\end{align}
The lower bound for $\yktn$ is always attained for $\yktzl$, the upper bound for $\yktzu$.
Note that when $s^k_t/n_k \in [\yktzl, \yktzu]$,
both the lower and the upper bound for $\yktn$ are attained using $\nktzu$,
this corresponds to the spotlight shape.
When $s^k_t/n_k \not\in [\yktzl, \yktzu]$,
the banana shape indicates that
we will get one of the bounds for $\yktn$ using $\nktzl$.

@@

The different locations and sizes of $\PktN$ in the conflict versus the no conflict case are then, in turn,
also reflected in the corresponding sets of Beta cdfs and Beta-Binomial cmfs.
As an example, the sets corresponding to the parameter sets in Figure~\ref{fig:paramsets}
are depicted in Figure~\ref{fig:betaset-binomset}.

In the no conflict case (Posterior 1, top row),
the reduction of the $\yktn$ range in $\PktN$ leads to a much smaller set
of Beta and Beta-Binomial distributions.
E.g., the range of predictive probabilities
that two out of a set of five components of type $k$ function at time $t$ 
has changed from $[0.10, 0.28]$ to $[0.11,0.14]$.
This reflects the gain in precision due to test data in accordance with prior assumptions.

In the prior-data conflict case (Posterior 2, bottom row),
the wide $\yktn$ range in $\PktN$ leads to a set of Beta and Beta-Binomial distributions
that is much larger than in the no conflict case.
Here, the range of posterior predictive probabilities
that two out of a set of five components of type $k$ function at time $t$
is now $[0.86, 1.00]$, i.e., less precise than in the no conflict case.
Using sets of Beta priors, the resulting set of posterior predictive Beta-Binomial distributions
adequately reflects the precision of prior information,
the amount of data, and prior-data conflict.

Furthermore, with sets of Beta priors it is also possible
to express prior ignorance in an appropriate way,
by letting $\yktzl \to 0$ and $\yktzu \to 1$
for some or all $t \in {\cal T}$.
(It is not advisable to choose $\yktzl = 0$ and $\yktzu = 1$,
as this can lead to improper posterior predictive distributions.
E.g., for any $t < \min(\vec{t}^k)$,
we would get $\yktnu = 1$, leading to one argument of the Beta function
in the denominator of \eqref{eq:postpredCny} being zero.)
Choosing these limits for $\yktz$ means that we are prepared to give only trivial bounds for the functioning probability,
and do not wish to make any commitments about $p^k_t$ a priori.
Compare this to the uniform prior over $[0,1]$, a usual choice of `noninformative' prior,
being a Beta prior with $\alpha^k_t = \beta^k_t = 1$, i.e., $\nktz = 2$, $\yktz = 0.5$.
Taking this for all $t \in {\cal T}$ means that we expect the component reliability function
to be on average 1/2 for all $t$, a very peculiar assumption,
and, in our view, quite incompatible with the notion of prior ignorance.

%With the bounded domain for $\yktz$ 
%\td{***called near-noninformative, for near-noninformative sets when $\yz$ is not bounded see
%benavoli zaffalon papers***}
%
%As a note regarding the bounds for $\nktz$,
%for an informative prior the choice of $\nktzl$ is relevant for posterior inferences in case of prior-data conflict,
%as then one of the bounds for $\ykt$ is attained for $\nktzl$.
%In a near-noninformative setting, the choice of $\nktzl$ is instead not relevant,
In a near-noninformative setting, the choice of $\nktzl$ is not relevant,
as due to \eqref{eq:ysetupdate}, both lower and upper bound for $\yktn$ are obtained with $\nktzu$; 
indeed $\yktzl > 0$ and $\yktzu < 1$ can be chosen such that
$\frac{s^k_t}{n_k} \in [\yktzl, \yktzu]$ for all $t \in \big(\min(\vec{t}^k), \max(\vec{t}^k)\big)$.
Obviously, one cannot have prior-data conflict in case of prior near ignorance.

>>>>

As was shown by \citet{2009:WalterAugustin}, %Walter \& Augustin (2009),
we can have both tractability and meaningful reaction to prior-data conflict
by using sets of priors $\MktZ$ produced by parameter sets $\PktZ = [\nktzl, \nktzu] \times [\yktzl, \yktzu]$
(a detailed discussion of different choices for $\PktZ$ is given in \citet[\S 3.1]{2013:diss-gw}.)
In our model, each prior parameter pair $(\nktz, \yktz) \in \PktZ$
corresponds to a Beta prior, thus $\MktZ$ is a set of Beta priors.
The set of posteriors $\MktN$ is obtained by updating each prior in $\MktZ$ according to Bayes' Rule.
This element-by-element updating can be rigorously justified
as ensuring coherence \citep[\S 2.5]{1991:walley}, and was termed ``Generalized Bayes' Rule'' by \citet[\S 6.4]{1991:walley}.
Due to conjugacy, $\MktN$ is a set of Beta distributions with parameters $(\nktn, \yktn)$,
obtained by updating $(\nktz, \yktz) \in \PktZ$ according to \eqref{eq:nyupdate},
leading to the set of updated parameters
\begin{align}
\PktN &= \Big\{ (\nktn, \yktn) \mid (\nktz, \yktz) \in \PktZ = [\nktzl, \nktzu] \times [\yktzl, \yktzu] \Big\}\,.
\label{eq:paramsets}
\end{align}

Examples for parameter sets $\PktZ$ and $\PktN$ as in \eqref{eq:paramsets} are depicted in Figure~\ref{fig:paramsets}.
Such rectangular prior parameter sets $\PktZ$ have been shown
to balance desirable model properties and ease of elicitation 
(see \citet[pp.~123f]{2013:diss-gw} or \citet{Troffaes2013a-short}).
%
For each component type $k$ and time point $t$,
one need only specify the four parameters $\nktzl, \nktzu, \yktzl, \yktzu$
(so in total $4 \times |{\cal T}|$ parameters are needed to define the set of prior distributions
on the survival function of each component.)
%notwithstanding this, the model has favourable inference properties.

@@

A desirable inference property arising from this setup is that
the posterior parameter set $\PktN$ is not rectangular in the way that the prior parameter set is.
Indeed, the shape of $\PktN$ depends on the presence or absence of prior-data conflict,
which is naturally operationalised as $s^k_t/n_k \not\in [\yktzl, \yktzu]$:
that is, prior-data conflict is defined to occur when, at time $t$, the observed fraction of functioning components
is outside its \emph{a priori} expected range.

First, in the absence of prior-data conflict, 
$\PktN$ shrinks in the $\ykt$ dimension;
how much it shrinks depending on $\nktz \in [\nktzl, \nktzu]$,
leading to the so-called spotlight shape depicted in Figure~\ref{fig:paramsets} (left).
Since $\yktn$ gives the posterior expectation for the functioning probability $p_t^k$,
shorter $\yktn$ intervals mean more precise knowledge about $p_t^k$.
Also, the variance interval for $p_t^k$ (not shown) will shorten and shift towards zero,
as the Beta distributions in $\MktN$ will be more concentrated
due to the increase of $\nktz$ to $\nktn$.

Alternatively, when there is conflict between prior and observed data (i.e.\ $s^k_t/n_k \not\in [\yktzl, \yktzu]$),
$\PktN$ instead adopts the so-called `banana shape',
arising from the intervals for $\yktn$  being shifted closer to $s^k_t/n_k$
for lower $\nktn$ values than for higher $\nktn$ values, see Figure~\ref{fig:paramsets} (right).
Overall, this results in a wider $\yktn$ interval compared to the no conflict case, 
reflecting the extra uncertainty due to prior-data conflict.  
In other words, the posterior sets make more cautious probability statements about $p_t^k$, as desired in this scenario.

Based on these shapes and \eqref{eq:nyupdate}, it is possible to deduce the following expressions
for the lower and upper bounds of $\yktn$:
\begin{align}
\begin{aligned}
\min_{\PktN} \yktn &=
 \begin{cases}
 %\dfrac{\nktzu \yktzl + s^k_t}{\nktzu + n_k} & \text{if } s^k_t/n_k \ge \yktzl \\
 \big(\nktzu \yktzl + s^k_t\big) / \big(\nktzu + n_k\big) & \text{if } s^k_t/n_k \ge \yktzl \\
 %\dfrac{\nktzl \yktzl + s^k_t}{\nktzl + n_k} & \text{if } s^k_t/n_k <   \yktzl
 \big(\nktzl \yktzl + s^k_t\big) / \big(\nktzl + n_k\big) & \text{if } s^k_t/n_k <   \yktzl
 \end{cases}\,,\\
\max_{\PktN} \yktn &=
 \begin{cases}
 %\dfrac{\nktzu \yktzu + s^k_t}{\nktzu + n_k} & \text{if } s^k_t/n_k \le \yktzu \\
 \big(\nktzu \yktzu + s^k_t\big) / \big(\nktzu + n_k\big) & \text{if } s^k_t/n_k \le \yktzu \\
 %\dfrac{\nktzl \yktzu + s^k_t}{\nktzl + n_k} & \text{if } s^k_t/n_k >   \yktzu
 \big(\nktzl \yktzu + s^k_t\big) / \big(\nktzl + n_k\big) & \text{if } s^k_t/n_k >   \yktzu
 \end{cases}\,.
\end{aligned}
\label{eq:ysetupdate}
\end{align}
Note that the lower bound for $\yktn$ is always attained at $\yktzl$, the upper bound at $\yktzu$.
Also note that when $s^k_t/n_k \in [\yktzl, \yktzu]$,
both the lower and the upper bounds for $\yktn$ are attained at $\nktzu$,
corresponding to the spotlight shape.
However, when $s^k_t/n_k \not\in [\yktzl, \yktzu]$,
the banana shape indicates that
one of the bounds for $\yktn$ is attained at $\nktzl$.

@@

The different locations and sizes of $\PktN$ in the conflict versus no conflict case are then, in turn,
also reflected in the corresponding sets of Beta cdfs and Beta-Binomial cmfs.
As an example, those corresponding to the parameter sets in Figure~\ref{fig:paramsets}
are depicted in Figure~\ref{fig:betaset-binomset}.

In the no conflict case (Posterior 1, top row),
the reduction of the $\yktn$ range in $\PktN$ leads to a much smaller set
of Beta and Beta-Binomial distributions.
For example, the range of predictive probabilities
that two out of a set of five components of type $k$ function at time $t$ 
has changed from $[0.10, 0.28]$ \emph{a priori} to $[0.11,0.14]$ \emph{a posteriori}.
This reflects the gain in precision due to test data in accordance with prior assumptions.

In contrast, for the prior-data conflict case (Posterior 2, bottom row),
the wide $\yktn$ range in $\PktN$ leads to a set of Beta and Beta-Binomial distributions
that is much larger than in the no conflict case.
Here, the range of posterior predictive probabilities
that two out of a set of five components of type $k$ function at time $t$
is now $[0.86, 1.00]$ \emph{a posteriori}, i.e., less precise than in the no conflict case.
Using sets of Beta priors, the resulting set of posterior predictive Beta-Binomial distributions
reflects the precision of prior information,
the amount of data, and prior-data conflict.

Furthermore, with sets of Beta priors it is also possible
to express prior ignorance
by letting $\yktzl \to 0$ and $\yktzu \to 1$
for some or all $t \in {\cal T}$.
(Note that it is not advisable to choose $\yktzl = 0$ and $\yktzu = 1$,
as this can lead to improper posterior predictive distributions.
For example, at any $t < \min(\vec{t}^k)$,
we would have $\yktnu = 1$, leading to one argument of the Beta function
in the denominator of \eqref{eq:postpredCny} being zero.)
These limits for $\yktz$ imply we are only prepared to give trivial bounds for the functioning probability
and do not wish to commit to any specific knowledge about $p^k_t$ \emph{a priori}.
This provides a more natural choice of `noninformative' prior over $[0,1]$ than the usual choice of
a Beta prior with $\alpha^k_t = \beta^k_t = 1$ (or $\nktz = 2$, $\yktz = 0.5$).
Such a prior for all $t \in {\cal T}$ actually reflects a belief that the component reliability function
is on average 1/2 for all $t$, which is not an expression of ignorance, but rather a very specific (and arguably peculiar) prior belief.

%With the bounded domain for $\yktz$ 
%\td{***called near-noninformative, for near-noninformative sets when $\yz$ is not bounded see
%benavoli zaffalon papers***}
%
%As a note regarding the bounds for $\nktz$,
%for an informative prior the choice of $\nktzl$ is relevant for posterior inferences in case of prior-data conflict,
%as then one of the bounds for $\ykt$ is attained for $\nktzl$.
%In a near-noninformative setting, the choice of $\nktzl$ is instead not relevant,
In a near-noninformative setting, the choice of $\nktzl$ is not relevant,
because \eqref{eq:ysetupdate} implies both lower and upper bound for $\yktn$ are obtained with $\nktzu$.
In particular, $\yktzl > 0$ and $\yktzu < 1$ can be chosen such that
$\frac{s^k_t}{n_k} \in [\yktzl, \yktzu]$ for all $t \in \big(\min(\vec{t}^k), \max(\vec{t}^k)\big)$.
Naturally, one cannot have prior-data conflict in cases of near prior ignorance.
